---
title: "US Doctorate Estimates"
format: pdf
editor: visual
date: October 03 2024
date-format: long
number-sections: true
bibliography: references.bib
---

## Introduction

In this paper we explore the US 2022 Census Data, sourced from @ipums2024 We use @citeR, @readr, @dplyr, @xie2023knitr

## Running Code

```{r}
#| echo: false
#| message: false
library(readr)
library(dplyr)
library(knitr)
```

The data is obtained from IPSUMS USA. Once on their website, navigate to Get Data. This brings up a search function. First, click select samples, deselect the "Default sample from each year" tick, and manually tick 2022 ACS. This is data taken in the US, not quite at the scale of a census, but involves much more thorough questioning. Click on 'Submit Summary' to add this dataset. Under Harmonized Variables, select the following through the dropdown menus: Household -> State -> STATEICP, Person -> Demographic -> SEX, Person -> Education -> EDUC. Finally, click on 'View Cart' and follow the steps required to download data (an IPSUMS account is required when checking out).

```{r}
#| echo: false
#| message: false
usa_census_data <- read_csv(here::here("data/usa_00003.csv"))
colnames(usa_census_data)

# Step 1: Calculate the number of respondents with a doctoral degree (EDUC == 116) in each state
doctorate_data <- usa_census_data %>%
  filter(EDUCD == 116) %>%  # Filter for doctoral degree respondents (code 116)
  group_by(STATEICP) %>%   # Group by state
  summarise(doctoral_count = n())  # Count respondents with doctoral degrees in each state

actual_total <- usa_census_data %>%
  group_by(STATEICP) %>%
  summarise(actual_total_respondents = n())

```

The ratio estimator approach involves the ratio of two random variables (R = a_x/a_y. It is used to estimate the population given a ratio and a preexisting population value (a_y = R*a_x). In this case, we are using California's ratio of doctorates (R), and the number of correspondents in California (a_x) to estimate the number of correspondents in other states (a_y).

```{r}
#| echo: false
#| message: false
# Step 2: Apply the ratio estimators approach
# Total number of respondents in California (across all education levels)
total_california_respondents <- 391171

# Get the number of respondents with doctoral degrees in California
doctoral_california <- doctorate_data %>% filter(STATEICP == 71) %>% pull(doctoral_count)

# Calculate the ratio r = (doctoral degree respondents in California) / (total respondents in California)
ratio <- doctoral_california / total_california_respondents

# Step 3: Estimate the total number of respondents in each state using this ratio
# For other states, estimate total respondents using the same ratio
estimated_total <- doctorate_data %>%
  mutate(estimated_total_respondents = doctoral_count / ratio)

combined_data <- left_join(actual_total, estimated_total, by = "STATEICP")

# View the results (estimated total respondents in each state based on the ratio)
knitr::kable(combined_data, caption = "Comparison of Actual and Estimated Total Respondents in Each State Based on Doctoral Degree Ratio")

```

Our estimated number of respondents were much higher than the actual number of respondents for most states that are not California. One reason could be that California's count of doctorates is much larger than most other states, which would inflate the correspondent estimator. If we had used another state for our estimator, one closer to the average number across all states, the estimator would get closer to the actual number.

## References